---
title: "spot_clustering_2Ch_2matches"
author: "Heesoo Song"
date: '2022 12 8 '
output: html_document
---

```{r}
library(cluster)
library(factoextra) 
library(dplyr)
library(ggplot2)
library(reshape2)
library(VennDiagram)
library(RColorBrewer)
library(plyr)
```


# 0. Load data
```{r}
base = "C:/Users/pc/Desktop/UNIST_internship/Sample_Image/Positive/PB417_01/"
#base = "C:/Users/pc/Desktop/UNIST_internship/Sample_Image/Negative/2/"
vis_path = paste0(base, "analysis_20221207/")

#file = "Spot_matching_result_imputated.csv"
file = "Spot_matching_result.csv"

# Two options for intensity value: "int" or "int_sig"
intensity = "int_sig"

# Channels of interest
C1 = FALSE
C2 = TRUE
C3 = FALSE
C4 = TRUE

id = c("C1_id", "C2_id", "C3_id", "C4_id")
channels = c("C1", "C2", "C3", "C4")
sigma = c("C1_sigma", "C2_sigma", "C3_sigma", "C4_sigma")
int = c("C1_int", "C2_int", "C3_int", "C4_int")
int_sig = c("C1_int_sig", "C2_int_sig", "C3_int_sig", "C4_int_sig")

id = id[c(C1, C2, C3, C4)]
channels = channels[c(C1, C2, C3, C4)]
sigma = sigma[c(C1, C2, C3, C4)]
int = int[c(C1, C2, C3, C4)]
int_sig = int_sig[c(C1, C2, C3, C4)]
```

```{r}
df_match <- read.csv(paste0(base, file))
print(colnames(df_match))
```

## 0-1. Limit Ranges
```{r}
# 1. Limit sigma (30<sigma<250)
if (C1){
  index_C1 <- which(df_match$C1_sigma < 42.6 | df_match$C1_sigma > 250)
  df_match$C1_sigma[index_C1] <- NA
  df_match$C1_int_sig[index_C1] <- NA
}
if (C2){
  index_C2 <- which(df_match$C2_sigma < 42.6 | df_match$C2_sigma > 250)
  df_match$C2_sigma[index_C2] <- NA
  df_match$C2_int_sig[index_C2] <- NA
} 
if (C3){
  index_C3 <- which(df_match$C3_sigma < 42.6 | df_match$C3_sigma > 250)
  df_match$C3_sigma[index_C3] <- NA
  df_match$C3_int_sig[index_C3] <- NA
}
if (C4){
  index_C4 <- which(df_match$C4_sigma < 42.6 | df_match$C4_sigma > 250)
  df_match$C4_sigma[index_C4] <- NA
  df_match$C4_int_sig[index_C4] <- NA
}
```

```{r}
# 2. Limit intensity (intensity > 1000)
if (intensity == "int"){
  if (C1){
    index_C1 <- which(df_match$C1_int < 1000)
    df_match$C1_sigma[index_C1] <- 0
    df_match$C1_int[index_C1] <- 0
  }
  if (C2){
    index_C2 <- which(df_match$C2_int < 1000)
    df_match$C2_sigma[index_C2] <- NA
    df_match$C2_int[index_C2] <- NA
  }
  if (C3){
    index_C3 <- which(df_match$C3_int < 1000)
    df_match$C3_sigma[index_C3] <- NA
    df_match$C3_int[index_C3] <- NA
  }
  if (C4){
    index_C4 <- which(df_match$C4_int < 1000)
    df_match$C4_sigma[index_C4] <- 0
    df_match$C4_int[index_C4] <- 0
  }
  
} #else if (intensity == "int_sig"){
#   if (C1){
#     index_C1 <- which(df_match$C1_int_sig < 1000)
#     df_match$C1_sigma[index_C1] <- 0
#     df_match$C1_int[index_C1] <- 0
#   }
#   if (C2){
#     index_C2 <- which(df_match$C2_int_sig < 1000)
#     df_match$C2_sigma[index_C2] <- 0
#     df_match$C2_int[index_C2] <- 0
#   }
#   if (C3){
#     index_C3 <- which(df_match$C3_int_sig < 1000)
#     df_match$C3_sigma[index_C3] <- 0
#     df_match$C3_int[index_C3] <- 0
#   }
#   if (C4){
#     index_C4 <- which(df_match$C4_int_sig < 1000)
#     df_match$C4_sigma[index_C4] <- 0
#     df_match$C4_int[index_C4] <- 0
#   }
# }
```

```{r}
# 3. Remove positions that the spot is out of range in all channels
df_match <- df_match[!(is.na(df_match[,sigma[1]]) | is.na(df_match[,sigma[2]])),]

if (intensity == "int"){
  df_meta <- df_match[,c(sigma, int)]
} else if (intensity == "int_sig"){
  df_meta <- df_match[,c(sigma, int_sig)]
}
```



# 1. Explore Data

## 1-1. Pair plot
```{r}
#pdf(file = paste0(vis_path, "pairs_plot.pdf"), width = 7, height = 7)
pairs(df_meta)
#dev.off()
```


## 1-2. Venn Diagram
```{r, include=FALSE}
C1_venn <- df_match[!is.na(df_match[,id[1]]), "X"]
C2_venn <- df_match[!is.na(df_match[,id[2]]), "X"]

# Chart
venndiagram <- venn.diagram(
  x = list(C1_venn, C2_venn),
  category.names = channels,
  filename = paste0(vis_path, 'venn_diagram.png'),
  disable.logging = TRUE,
  output = TRUE,
  
  # Output features
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 300,
  compression = "lzw",
  
  # Circles
  lwd = 2,
  lty = 'blank',
  fill = brewer.pal(8, "Set1")[c(1, 2)],
  
  # Numbers
  cex = .6,
  fontface = "bold",
  fontfamily = "sans",
  
  # Set names
  cat.cex = 0.6,
  cat.fontface = "bold",
  cat.default.pos = "outer",
  #cat.pos = c(-27, 27, 135),
  #cat.dist = c(0.055, 0.055, 0.085),
  cat.fontfamily = "sans",
  #rotation = 1
)
```

## 1-3. Number of spots in each channel
```{r}
n_spots <- data.frame(channel = channels, freq = c(length(C1_venn), length(C2_venn)))

#pdf(file = paste0(vis_path, "n_spots_each_channel.pdf"), width = 8, height = 6)
ggplot(n_spots, aes(x = channel, y = freq)) + 
  geom_bar(position = "dodge", stat="identity", width=1, color = "white") +
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (freq + 1000), label = freq, fontface = 2), size=5) +
  labs(title = "Number of Spots in Each Channel")
#dev.off()
```

## 1-4. Average intensity/sigma of spots in each channel
```{r}
if (intensity == "int"){
  avg_int_prep <- df_match[,int]
} else if (intensity == "int_sig"){
  avg_int_prep <- df_match[,int_sig]
}

avg_sig_prep <- df_match[,sigma]
avg_sig_prep[avg_sig_prep == 0] <- NA

avg_int <- apply(avg_int_prep, 2, mean, na.rm = TRUE)
avg_sig <- apply(avg_sig_prep, 2, mean, na.rm = TRUE)

df_avg_int <- data.frame(Channel = channels, Average_Intensity = avg_int)
df_avg_sig <- data.frame(Channel = channels, Average_Sigma = avg_sig)

df_avg_int$Average_Intensity <- round(df_avg_int$Average_Intensity,2)
df_avg_sig$Average_Sigma <- round(df_avg_sig$Average_Sigma,2)
df_avg_sig$Average_FWHM <- 2.35*df_avg_sig$Average_Sigma
```

```{r}
#pdf(file = paste0(vis_path, "avg_int_histogram.pdf"), width = 8, height = 6)
ggplot(df_avg_int, aes(x = Channel, y = Average_Intensity)) + 
  geom_bar(position = "dodge", stat="identity", width=1, color = "white") +
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (Average_Intensity + Average_Intensity*0.1), label = Average_Intensity, fontface = 2), size=5) +
  labs(title = "Average Intensity in Each Channel")
#dev.off()
```

```{r}
#pdf(file = paste0(vis_path, "avg_sig_histogram.pdf"), width = 8, height = 6)
ggplot(df_avg_sig, aes(x = Channel, y = Average_Sigma)) + 
  geom_bar(position = "dodge", stat="identity", width=1, color = "white") +
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (Average_Sigma + 3), label = Average_Sigma, fontface = 2), size=5) +
  labs(title = "Average Sigma in Each Channel")
#dev.off()
```

```{r}
#pdf(file = paste0(vis_path, "avg_sig_histogram.pdf"), width = 8, height = 6)
ggplot(df_avg_sig, aes(x = Channel, y = Average_FWHM)) + 
  geom_bar(position = "dodge", stat="identity", width=1, color = "white") +
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (Average_FWHM + 10), label = Average_FWHM, fontface = 2), size=5) +
  labs(title = "Average FWHM in Each Channel")
#dev.off()
```

## 1-5. Compare Intensity/Sigma distribution in channels
```{r}
multi_sig_dist <- avg_sig_prep %>% melt()
multi_sig_dist <- multi_sig_dist[which(multi_sig_dist$value > 0),]
cdat <- ddply(multi_sig_dist, "variable", summarise, rating.mean = mean(value))
#CI_low <- ddply(multi_sig_dist, "variable", summarise, CI = as.numeric(t.test(value, conf.level = 0.95)[4]$conf.int[1]))
#CI_high <- ddply(multi_sig_dist, "variable", summarise, CI = as.numeric(t.test(value, conf.level = 0.95)[4]$conf.int[2]))

#pdf(file = paste0(vis_path, "sigma_distribution_histogram.pdf"), width = 8, height = 6)
ggplot(multi_sig_dist, aes(x = value, fill = variable)) +
  geom_histogram(binwidth=5, alpha = .5, position = "identity")+
  geom_vline(data = cdat, aes(xintercept=rating.mean, colour = variable),linetype = "dashed", size = 1) +
  xlab("sigma")
  #geom_vline(data = CI_low, aes(xintercept=CI, colour = variable), size = 0.5) +
  #geom_vline(data = CI_high, aes(xintercept=CI, colour = variable), size = 0.5)
#dev.off()

#pdf(file = paste0(vis_path, "sigma_distribution_barplot.pdf"), width = 8, height = 6)
ggplot(multi_sig_dist, aes(x=variable, y=value, color=variable)) +
  geom_boxplot() +
  theme_bw()
#dev.off()
```

```{r}
multi_int_dist <- avg_int_prep %>% melt()
multi_int_dist <- multi_int_dist[c(rownames(multi_sig_dist)),]
cdat <- ddply(multi_int_dist, "variable", summarise, rating.mean = mean(value))

#pdf(file = paste0(vis_path, "intensity_distribution_histogram.pdf"), width = 8, height = 6)
ggplot(multi_int_dist, aes(x = value, fill = variable)) +
  geom_histogram(binwidth=100, alpha = .5, position = "identity")+
  geom_vline(data = cdat, aes(xintercept=rating.mean, colour = variable),linetype = "dashed", size = 1) +
  xlim(0, 10000) +
  xlab("intensity")
#dev.off()

#pdf(file = paste0(vis_path, "intensity_distribution_barplot.pdf"), width = 8, height = 6)
ggplot(multi_int_dist, aes(x=variable, y=value, color=variable)) +
  geom_boxplot() +
  theme_bw()
#dev.off()
```

# 2. Cluster Analysis

## 2-1. Normalization & Imputation
```{r}
# Normalization
df_meta <- log10(df_meta)
df_meta[sapply(df_meta, is.infinite)] <- NA
means <- apply(df_meta,2,mean, na.rm=TRUE)
sds <- apply(df_meta,2,sd, na.rm=TRUE)
df_meta_norm <- scale(df_meta,center=means,scale=sds) %>% data.frame()

# 1) Imputate NA with normalized zero(sigma = 0.1, intensity = 0.1)
df_meta_norm_imputate <- (log10(0.1) - means)/sds
for (i in 1:length(df_meta_norm_imputate)){
  df_meta_norm[is.na(df_meta_norm[,i]),i] <- df_meta_norm_imputate[i]
}

# 2) Impuate NA with mean
#df_meta_norm[sapply(df_meta_norm, is.na)] <- 0
```

## 2-2. Histogram: Sigma/intensity distribution after normalization
```{r}
df_meta_norm_sig <- df_meta_norm[,sigma]
df_meta_norm_sig_dist <- df_meta_norm_sig %>% melt()
#df_meta_norm_sig_dist <- df_meta_norm_sig_dist[,c(2,3)]
colnames(df_meta_norm_sig_dist)[1] <- "variable"
df_meta_norm_sig_dist <- df_meta_norm_sig_dist[c(rownames(multi_sig_dist)),]
cdat <- ddply(df_meta_norm_sig_dist, "variable", summarise, rating.mean = mean(value))

#pdf(file = paste0(vis_path, "sigma_distribution_norm_histogram.pdf"), width = 8, height = 6)
ggplot(df_meta_norm_sig_dist, aes(x = value, fill = variable)) +
  geom_histogram(binwidth=.1, alpha = .5, position = "identity")+
  geom_vline(data = cdat, aes(xintercept=rating.mean, colour = variable),linetype = "dashed", size = 1) +
  xlab("sigma")
#dev.off()

#pdf(file = paste0(vis_path, "sigma_distribution_norm_barplot.pdf"), width = 8, height = 6)
ggplot(df_meta_norm_sig_dist, aes(x=variable, y=value, color=variable)) +
  geom_boxplot() +
  theme_bw()
#dev.off()
```

```{r}
if (intensity == "int"){
  df_meta_norm_int <- df_meta_norm[,int]
} else if (intensity == "int_sig"){
  df_meta_norm_int <- df_meta_norm[,int_sig]
}

df_meta_norm_int_dist <- df_meta_norm_int %>% melt()
#df_meta_norm_int_dist <- df_meta_norm_int_dist[,c(2,3)]
colnames(df_meta_norm_int_dist)[1] <- "variable"
df_meta_norm_int_dist <- df_meta_norm_int_dist[c(rownames(multi_int_dist)),]
cdat <- ddply(df_meta_norm_int_dist, "variable", summarise, rating.mean = mean(value))

#pdf(file = paste0(vis_path, "intensity_distribution_norm_histogram.pdf"), width = 8, height = 6)
ggplot(df_meta_norm_int_dist, aes(x = value, fill = variable)) +
  geom_histogram(binwidth=.1, alpha = .5, position = "identity")+
  geom_vline(data = cdat, aes(xintercept=rating.mean, colour = variable),linetype = "dashed", size = 1) +
  xlab("intensity")
#dev.off()

#pdf(file = paste0(vis_path, "intensity_distribution_norm_barplot.pdf"), width = 8, height = 6)
ggplot(df_meta_norm_int_dist, aes(x=variable, y=value, color=variable)) +
  geom_boxplot() +
  theme_bw() 
#dev.off()
```

## 2-3. Calculate distance and determine cluster number
```{r}
# # Euclidean distance: observation with high values of features will be clustered together
# # Correlation based distance: Identify clusters of observation regardless of their magnitude
# distance = dist(df_meta_norm)
# df_meta.hclust = hclust(distance, method = "ward.D2")
# 
# #pdf(file = paste0(vis_path, "dendrogram.pdf"), width = 10, height = 7)
# plot(df_meta.hclust,hang=-1, labels = FALSE, main='Hierarchical Cluster')
# #dev.off()
```

## 2-4. Determine the best number of clusters
```{r}
# sil_coef_vector <- c()
# for (num_cluster in 2:10){
#   print(paste0("computing cluster number ", num_cluster))
#   sil_cutree <- cutree(df_meta.hclust, k = num_cluster)
#   sil_cl <- silhouette(sil_cutree ,distance)
#   sil_cl_summary <- summary(sil_cl)
#   sil_coefficient <- as.vector(sil_cl_summary$si.summary["Mean"])
#   sil_coef_vector<- c(sil_coef_vector, sil_coefficient)
# }
# 
# #pdf(file = paste0(vis_path, "silhouette coefficient.pdf"), width = 10, height = 7)
# plot(2:10,sil_coef_vector,"b", xlab="k", ylab="silhouette coefficient")
# #dev.off()
# 
#pdf(file = paste0(vis_path, "average_silhouette coefficient.pdf"), width = 10, height = 7)
fviz_nbclust(df_meta_norm, kmeans, method = "silhouette")
#dev.off()
```

## 2-5. Clustering
```{r}
# K-means clustering
n_cluster = 3
k2 <- kmeans(df_meta_norm, centers = n_cluster, nstart = 25)  # centers = number of clusters
str(k2)

#pdf(file = paste0(vis_path, "clusters.pdf"), width = 10, height = 10)
fviz_cluster(k2, data = df_meta_norm, geom = "point", ggtheme = theme_bw())
#dev.off()

# # Fuzzy c-means clustering
# 
# n_cluster = 2
# k2.fcm <- fcm(df_meta_norm, centers=n_cluster, nstart=5)
# plotcluster(k2.fcm, cp=1, trans=TRUE)
# 
# k2 <- ppclust2(k2.fcm, "kmeans")
# fviz_cluster(k2, data = df_meta_norm, 
#                          ellipse.type = "convex",
#                          palette = "jco",
#                          geom = "point")
```

## 2-6. Pair plot indicated with clusters
```{r, message=FALSE, warning=FALSE}
df_match[,'cluster'] <- factor(k2$cluster)
df_meta_norm[,'cluster'] <- factor(k2$cluster)
```

```{r, warning = FALSE}
library(GGally)
#pdf(file = paste0(vis_path, "clusters_pairs.pdf"), width = 10, height = 10)

if (intensity == "int"){
 ggpairs(df_meta_norm, columns=c(sigma, int), aes(colour=cluster, alpha = 0.5), lower=list(continuous='points'), axisLabels='none')#, upper=list(continuous='blank')) 
} else if (intensity == "int_sig"){
  ggpairs(df_meta_norm, columns=c(sigma, int_sig), aes(colour=cluster, alpha = 0.5), lower=list(continuous='points'), axisLabels='none')#, upper=list(continuous='blank'))
}

#dev.off()
```

## 2-7. Save cluster result
```{r}
write.csv(df_match, file = paste0(base, "spot_matching_result_imputated_cluster.csv"))
```


# 3. Evaluate Clusters

## 3-1. Number of spots for each cluster
```{r}
cluster_histogram <- k2$cluster %>% table()
cluster_histogram_df <- data.frame(cluster = names(cluster_histogram), freq = as.numeric(cluster_histogram))

#pdf(file = paste(vis_path, "Cluster_freq_histogram.pdf", sep=""), width = 9, height = 7)
ggplot_cluster_histogram <- ggplot(cluster_histogram_df, aes(x=cluster, y=freq)) +
  geom_bar(stat="identity", width=1, color = "white") +
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (freq+10), label = freq, fontface = 2), size=8) +
  labs(title = "Number of Spots in Each Cluster")

print(ggplot_cluster_histogram)
#dev.off()
```

## 3-2. the proportion of cluster in each channel
```{r}
for (ch in channels){
  #pdf(file = paste(vis_path, paste0("cluster_proportion_in_channel_", ch, ".pdf"), sep=""), width = 9, height = 7)
  
  cluster <- df_match[!is.na(df_match[,paste0(ch, "_id")]),"cluster"]
  cluster_freq <- cluster %>% table()
  cluster_freq_df <- data.frame(cluster = names(cluster_freq), freq = as.numeric(cluster_freq))
  empty_cluster <- setdiff(c(1:n_cluster), names(cluster_freq))
  if (length(empty_cluster) > 0){
    for (i in 1:length(empty_cluster)){
      cluster_freq_df[(length(names(cluster_freq))+i),"cluster"] <- empty_cluster[i]
      cluster_freq_df[(length(names(cluster_freq))+i),"freq"] <- 0
      cluster_freq_df$cluster <- factor(cluster_freq_df$cluster, levels = c(1:n_cluster))
    }
  }
  
  ggplot_cluster_proportion <- ggplot(cluster_freq_df, aes(x=cluster, y=freq)) +
    geom_bar(stat="identity", width=1, color = "white") +
    #scale_fill_manual(values = c("1" = "#F8766D", "2" = "#D39200", "3" = "#00BA38", "4" = "#00aedb", "5" = "#C77CFF")) + #"5" = "#00B9E3", , "7" = "#FF61C3"
    theme_bw() +
    theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
    geom_text(aes(y = (freq+10), label = freq, fontface = 2), size=8) +
    labs(title = paste0("Cluster Proportion in channel ", ch))
  
  print(ggplot_cluster_proportion)
  #dev.off()
}

```

## 3-3. the proportion of cluster in each n_match
```{r}
#pdf(file = paste(vis_path, "cluster_proportion_in_n_match.pdf", sep=""), width = 9, height = 7)

n_match <- df_match[which(df_match$n_match > 1),c("cluster", "n_match")]
n_match_freq_df <- n_match %>% table() %>% melt() %>% data.frame()

n_match_freq_df$cluster <- factor(n_match_freq_df$cluster)
n_match_freq_df$n_match <- factor(n_match_freq_df$n_match)

ggplot_n_match_proportion <- ggplot(n_match_freq_df, aes(x=cluster, y=value, fill=cluster)) +
  geom_bar(position = "dodge", stat="identity", width=1, color = "white") +
  facet_wrap(~n_match) +
  scale_fill_manual(values = c("1" = "#F8766D", "2" = "#D39200", "3" = "#00BA38", "4" = "#00aedb", "5" = "#C77CFF")) + #"5" = "#00B9E3", , "7" = "#FF61C3"
  theme_bw() +
  theme(plot.title = element_text(size=20, hjust=0.5, face = "bold")) +
  geom_text(aes(y = (value + 7), label = value, fontface = 2), size=3) +
  labs(title = "Cluster Propotion in n_match")

print(ggplot_n_match_proportion)
#dev.off()
```

## 3-4. Mean sigma & intensity values in each cluster
```{r}
mean_df <- data.frame(C1_sigma = c(), C2_sigma = c(), C3_sigma = c(), C4_sigma = c(), C1_int = c(), C2_int = c(), C3_int = c(), C4_int = c())
for (c in 1:n_cluster){
  sigma <- df_match[which(df_match$cluster == c),c("C1_sigma", "C2_sigma", "C3_sigma", "C4_sigma")]
  int <- df_match[which(df_match$cluster == c), c("C1_int", "C2_int", "C3_int", "C4_int")]
  
  mean_sigma <- apply(sigma, 2, mean, na.rm=TRUE)
  mean_int <- apply(int, 2, mean, na.rm=TRUE)
  
  for (d in 1:4){
    mean_df[c, names(mean_sigma)[d]] <- mean_sigma[d]
    mean_df[c, names(mean_int)[d]] <- mean_int[d]
  }
}
mean_df <- mean_df[,c(1,3,5,7,2,4,6,8)]
mean_df

```

